Dominance Regions for Rank Two Cluster Algebras

### 2022-04-05 dominance_regions.tex
More reorganization and edits

\section{$\bfg$-vector Mutations}
  We begin by studying transformations of $\RR^2$ which determine the change of $\bfg$-vectors under mutations.
  Write $\phi_0:\RR^2\to\RR^2$ for the identity map and define piecewise-linear maps $\phi_{\pm 1}:\RR^2\to\RR^2$ as follows:
  \begin{equation}
    \phi_1(\lambda)
    =
    \begin{cases} 
      (-\lambda_1,c\lambda_1+\lambda_2) & \text{if $\lambda_1 \ge 0$;}\\
      (-\lambda_1,\lambda_2) & \text{if $\lambda_1 < 0$;}
    \end{cases}
    \qquad
    \phi_{-1}(\lambda)
    =
    \begin{cases} 
      (\lambda_1,-\lambda_2) & \text{if $\lambda_2 \ge 0$;}\\
      (\lambda_1+b\lambda_2,-\lambda_2) & \text{if $\lambda_2 < 0$.}
    \end{cases}
  \end{equation}
  For $k\in\ZZ$ with $|k|>1$, define piecewise-linear maps
  \[\phi_k
    :=
    \begin{cases}
      (\phi_{-1}^{-1}\phi_1)^j & \text{if $k=2j$;}\\
      \phi_1(\phi_{-1}^{-1}\phi_1)^j & \text{if $k=2j+1$.}\\
    \end{cases}
  \]
  These determine the dominance region as follows.
  \begin{definition}
    For $\lambda\in\ZZ^2$, define cones 
    \[\cC(\lambda):=\{(\lambda_1-r,\lambda_2+s):r,s\in\RR_{\ge0}\}\quad\text{ and }\quad \cC'(\lambda):=\{(\lambda_1+r,\lambda_2-s):r,s\in\RR_{\ge0}\}.\]
    The \emph{dominance region} $\cP(\lambda)$ is the intersection of $\bigcap_{j\in\ZZ}\phi_{2j}^{-1}\cC(\phi_{2j}\lambda)$ and $\bigcap_{j\in\ZZ} \phi_{2j+1}^{-1}\cC'(\phi_{2j+1}\lambda)$.
  \end{definition}

@@

  Next we aim to understand how inequalities transform under the action of a linear map, this will be the key tool for describing the dominance regions $\cP(\lambda)$ for $\lambda\in\cI$.
  \begin{lemma}
    \label{le:transformed inequalities}
    Let $A=(a_{ij})$ be an invertible $2\times 2$ matrix.
    Under the action of $A$, the region inside $\RR^2$ defined by the inequality $rx+sy\le t$ is transformed to the region defined by the inequality \[\frac{ra_{22}-sa_{21}}{a_{11}a_{22}-a_{12}a_{21}}x + \frac{-ra_{12}+sa_{11}}{a_{11}a_{22}-a_{12}a_{21}}y\le t.\]
  \end{lemma}
  \begin{proof}
    Assume $r,s,t\ne0$, the cases when one of these vanishes follows by the continuity of the transformation.
    Observe that the points $\big(\frac{t}{r},0\big)$ and $\big(0,\frac{t}{s}\big)$ lie on the boundary line of the region defined by the inequality $rx+sy\le t$.
    These are transformed under the action of $A$ into the points $\big(a_{11}\frac{t}{r},a_{21}\frac{t}{r}\big)$ and $\big(a_{12}\frac{t}{s},a_{22}\frac{t}{s}\big)$, respectively, which one readily checks lie along the purported boundary of the image region.
    By the linearity of the transformation of $\RR^2$ given by the matrix $A$, we see that either both the domain region and image region contain the origin or both do not, in particular this implies that the inequalities go in the same direction.
  \end{proof}

@@

\section{Lattice Mutations and Linear Inequalities}
  Let $B=(b_{ij})$ be an $n\times n$ skew-symmetrizable integer matrix, i.e. there exists a diagonal integer matrix $D=\diag(d_1,\ldots,d_n)$ so that $DB$ is skew-symmetric, more precisely we have $d_i b_{ij}=-d_j b_{ji}$.
  Write $\Col(B)$ for the span of the columns of $B$ inside $\RR^n$ and $\Col^+(B)$ for their non-negative span.
  Define $\Col_\ZZ(B)$ and $\Col^+_\ZZ(B)$ analogously over the integers.

  The matrix $B$ is \emph{acyclic} if there is no sequence $i_1,\ldots,i_m,i_{m+1}=i_1$ so that $b_{i_\ell i_{\ell+1}}>0$ for $1\le\ell\le m$.
  This condition is naturally interpreted in terms of an associated quiver $Q_B$ with $n$ vertices and an arrow $i\to j$ whenever $b_{ji}>0$ (if we need to use representation theory at any point we'll have to change this to be $\gcd(b_{ij},b_{ji})$ arrows $i\to j$).
  Indeed, this is precisely the condition that $Q_B$ has no oriented cycles.
  In the case when $B$ is acyclic, there exists a labeling $\{k_1,\ldots,k_n\}=\{1,\ldots,n\}$ so that $\ell<\ell'$ implies $b_{k_{\ell'} k_\ell}\ge 0$.
  The sequence $(k_1,\ldots,k_n)$ (resp. $(k_n,\ldots,k_1)$) is then said to be \emph{source-adapted} (resp. \emph{sink-adapted}) since the condition implies $k_1$ is a source and $k_n$ is a sink in $Q_B$.
  %\sayDR{We need to pin down this convention.}

@@

  For $t\in\TT_n$, define the piecewise-linear automorphisms $\phi_t:\RR^n\to\RR^n$ by
  \begin{itemize}
    \item $\phi_{t_0}=\Id$;
    \item if $t,t'\in\TT_n$ are joined by an edge labeled $k$, then $\phi_{t'}=\phi^t_k \phi_t$.
  \end{itemize}

  For $\lambda\in\RR^n$, define the \emph{$B_t$-cone} $\cC_t(\lambda):=\lambda+B_t\cdot\RR_{\ge0}^n$ pointed at $\lambda$.
  \begin{definition}
    For $\lambda\in\RR^n$, define the \emph{dominance region} 
    \[ \cP(\lambda) = \bigcap_{t\in\TT_n} \phi_t^{-1} \cC_t(\phi_t \lambda). \]
    When $\mu\in\cP(\lambda)$, we say \emph{$\lambda$ dominates $\mu$}.
  \end{definition}
  Our goal is to understand the dominance polytopes $\cP(\lambda)$ for all $\lambda$.
  \begin{lemma}
    The sets $\cP(\lambda)$ are polytopes.
  \end{lemma}

@@

\section{The affine case}
  In this section we compare our main result with the construction of \cite{RSW19} in the case $b=c=2$.
  We show that the continuous family of bases constructed there is precisely the deformation family constructed by \cite{Qin}.

>>>>

\section{$\bfg$-vector Mutations}
  We begin by studying transformations of $\RR^2$ which determine the change of $\bfg$-vectors under mutations.
  Write $\phi_0:\RR^2\to\RR^2$ for the identity map and define piecewise-linear maps $\phi_{\pm 1}:\RR^2\to\RR^2$ as follows:
  \begin{equation}
    \phi_1(\lambda)
    =
    \begin{cases} 
      (-\lambda_1,c\lambda_1+\lambda_2) & \text{if $\lambda_1 \ge 0$;}\\
      (-\lambda_1,\lambda_2) & \text{if $\lambda_1 < 0$;}
    \end{cases}
    \qquad
    \phi_{-1}(\lambda)
    =
    \begin{cases} 
      (\lambda_1,-\lambda_2) & \text{if $\lambda_2 \ge 0$;}\\
      (\lambda_1+b\lambda_2,-\lambda_2) & \text{if $\lambda_2 < 0$.}
    \end{cases}
  \end{equation}
  For $k\in\ZZ$ with $|k|>1$, define piecewise-linear maps
  \[\phi_k
    :=
    \begin{cases}
      (\phi_{-1}^{-1}\phi_1)^j & \text{if $k=2j$;}\\
      \phi_1(\phi_{-1}^{-1}\phi_1)^j & \text{if $k=2j+1$.}\\
    \end{cases}
  \]
  These determine the dominance region as follows.
  \begin{definition}
    For $\lambda\in\ZZ^2$ and $k\in\ZZ$, define cones 
    \[\cC_k(\lambda)
      :=
      \begin{cases}
        \{(\lambda_1-r,\lambda_2+s):r,s\in\RR_{\ge0}\} & \text{if $k=2j$;}\\
        \{(\lambda_1+r,\lambda_2-s):r,s\in\RR_{\ge0}\} & \text{if $k=2j+1$.}
      \end{cases}
    \]
    The \emph{dominance region} $\cP(\lambda)$ is the intersection  $\bigcap_{k\in\ZZ}\phi_k^{-1}\cC_k(\phi_k\lambda)$.
    When $\mu\in\cP(\lambda)$, we say \emph{$\lambda$ dominates $\mu$}.
  \end{definition}

@@

  Next we aim to understand how inequalities transform under the action of a linear map, this will be the key tool for describing the dominance regions $\cP(\lambda)$ for $\lambda\in\cI$.

  \begin{lemma}
    \label{le:transformed inequalities}
    Let $A=(a_{ij})$ be an invertible $2\times 2$ matrix.
    Under the action of $A$, the region inside $\RR^2$ defined by the inequality $rx+sy\le t$ is transformed to the region defined by the inequality \[\frac{ra_{22}-sa_{21}}{a_{11}a_{22}-a_{12}a_{21}}x + \frac{-ra_{12}+sa_{11}}{a_{11}a_{22}-a_{12}a_{21}}y\le t.\]
  \end{lemma}
  \begin{proof}
    Assume $r,s,t\ne0$, the cases when one of these vanishes follows by the continuity of the transformation.
    Observe that the points $\big(\frac{t}{r},0\big)$ and $\big(0,\frac{t}{s}\big)$ lie on the boundary line of the region defined by the inequality $rx+sy\le t$.
    These are transformed under the action of $A$ into the points $\big(a_{11}\frac{t}{r},a_{21}\frac{t}{r}\big)$ and $\big(a_{12}\frac{t}{s},a_{22}\frac{t}{s}\big)$, respectively, which one readily checks lie along the purported boundary of the image region.
    By the linearity of the transformation of $\RR^2$ given by the matrix $A$, we see that either both the domain region and image region contain the origin or both do not, in particular this implies that the inequalities go in the same direction.
  \end{proof}

@@

\section{Lattice Mutations and Linear Inequalities}

@@

@@

\section{The affine case}
  \label{sec:affine}
  In this section we compare our main result with the construction of \cite{RSW19} in the case $b=c=2$.
  We show that the continuous family of bases constructed there is precisely the deformation family constructed by \cite{Qin}.
